---
title: "Topic 2: Text Data in R Assignment"
author: "Mia Forsline"
date: "4/6/2022"
output: html_document
---

# Set up: Load necessary packages 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warnings = FALSE,
                      messages = FALSE)

library(jsonlite) #convert results from API queries into R-friendly formats 
library(tidyverse) 
library(tidytext) #text data management and analysis
library(ggplot2) #plot word frequencies and publication dates
```

# 2. Connect to the New York Times API and send a query

- Query phrase: "bark beetle" 

```{r api, eval = FALSE}
#create an object called x with the results of our query ("bark beetle")
# the from JSON flatten the JSON object, then convert to a data frame

key = "zwtPmtpUJiMGPfGjXz8maQyr3cx6YFdS"

x <- fromJSON("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=bark+beetle&api-key=zwtPmtpUJiMGPfGjXz8maQyr3cx6YFdS", 
              flatten = TRUE) #the string following "key=" is your API key 

class(x) #note that x is a list 

#convert x to a dataframe 
x <- x %>% 
  data.frame()

#the dataframe is 10 x 33
dim(x)

#check the variable names in the dataframe 
names(x)
```

## Examine a piece of text

- the data object has a variable called "response.docs.snippet" that contains a short excerpt, or "snippet" from the article. 
- check the snippet to ensure that the query picked out articles containing the correct query word ("bark beetle")

```{r snippet, eval=FALSE}
x$response.docs.snippet[9]
```

# Set some parameters for a bigger query

- searching for articles from 1990 - 2022 

```{r}
term <- "bark+beetle" # Need to use + to string together separate words
begin_date <- "19900101" #start date: Jan 1, 1990
end_date <- "20220101" #end date: Jan 1, 2022
key = "zwtPmtpUJiMGPfGjXz8maQyr3cx6YFdS"

#construct the query url using API operators
baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term,
                  "&begin_date=",begin_date,
                  "&end_date=",end_date,
                  "&facet_filter=true&api-key=",key, sep="")

#examine our query url
baseurl
```

Obtain multiple pages of query results 

- retrieved 35 pages of results 
- retrieved 358 articles 

```{r, eval=FALSE}
initialQuery <- fromJSON(baseurl)

maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1) 

pages <- list()

for(i in 0:maxPages){
  nytSearch <- fromJSON(paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
  message("Retrieving page ", i)
  pages[[i+1]] <- nytSearch 
  Sys.sleep(6) 
}

class(nytSearch)
#need to bind the pages and create a tibble from nytDa

nytDat <- rbind_pages(pages)
```

Visualize the various types of materials that were retrieved 
```{r article-type}
#nytDat <- read.csv("nytDat.csv") # obtained from 
#dim(nytDat)

nytDat %>% 
  group_by(response.docs.type_of_material) %>%
  summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=response.docs.type_of_material, fill=response.docs.type_of_material), 
           stat = "identity") + 
  coord_flip() + 
  theme_classic() + 
  ylab("Material Type") + 
xlab("% of Materials Retrieved")
```
# 3. Visuzalize publications per day 

```{r date-plot}
nytDat %>%
  mutate(pubDay=gsub("T.*","",response.docs.pub_date)) %>%
  group_by(pubDay) %>%
  summarise(count=n()) %>%
  filter(count >= 2) %>%
  ggplot() +
  geom_bar(aes(x=reorder(pubDay, count), y=count), stat="identity") + coord_flip() + 
  labs(y = "Publication Day", 
       x = "Number of Articles Published")
```

The New York Times doesn't make full text of the articles available through the API. But we can use the first paragraph of each article.

- unnest the words from the first paragraph of each article 
```{r plot_frequencies}
names(nytDat)

paragraph <- names(nytDat)[6] #The 6th column, "response.doc.lead_paragraph", is the one we want here.  

tokenized <- nytDat %>%
  unnest_tokens(word, paragraph)
```

View the unique words pulled from the first paragraph of each article 
```{r plot_frequencies}
tokenized[,34] 
```

# 3. Visualize word frequency plotsusing the first paragraph of each article 
- note that there are so many words that the graph is extremely difficult to read 

```{r plot_frequencies}
tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

# 4. Since this visual is not very helpful, we transform the corpus in various ways to help filter out less meaningful words from our visual 

## Incorporate stop words 
```{r}
#call common stop words from a lexicon 
data(stop_words)
stop_words

#use an anti-join to remove the stop words from our tokenized object 
tokenized <- tokenized %>%
  anti_join(stop_words)
```

# Visualize the most common words (excluding stop words)
```{r}
# tokenized %>%
#   count(word, sort = TRUE) %>%
#   filter(n > 5) %>%
#   mutate(word = reorder(word, n)) %>%
#   ggplot(aes(n, word)) +
#   geom_col() +
#   labs(y = NULL)
```


```{r inspect, eval=FALSE}
#inspect the list of tokens (words)
tokenized$word
count(tokenized)
```

## Stem tribe words

- remove plurals and other varying forms of a core word 
```{r cleaning, eval=FALSE}
clean_tokens <- str_replace_all(tokenized$word,"fringe[a-z,A-Z]*","fringe")

#check to see if the transformation worked 
# %>% 
#   as.data.frame()
# clean_tokens$.
```

## Remove numbers 
```{r cleaning, eval=FALSE}
clean_tokens <- str_remove_all(clean_tokens, "[:digit:]") 

#check to see if the transformation worked 
# %>% 
#   as.data.frame()
# clean_tokens$.
```

## Remove plural words 
```{r cleaning, eval=FALSE}
clean_tokens <- gsub("’s", '', clean_tokens) %>% 
  as.data.frame()
clean_tokens$.
```

Put clean tokens into the `tokenized` dataframe 
```{r cleaning, eval=FALSE}
tokenized$clean <- clean_tokens

# tokenized %>%
#   count(clean, sort = TRUE) %>%
#  #illegible with all the words displayed
#   mutate(clean = reorder(clean, n)) %>%
#   ggplot(aes(n, clean)) +
#   geom_col() +
#   labs(y = NULL)
```

# Visualize after cleaning

- limit the graph to only include words that occur at least 20 times 

```{r cleaning, eval=FALSE}
#remove the empty strings
tib <-subset(tokenized, clean!="")
#reassign
tokenized <- tib

#visualize again
tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 20) %>% 
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = "Most common words", 
       x = "Word Occurrence Frequency") + 
  theme_classic()
```

## Assignment (Due by Week 3)

1.  Create a free New York Times account (<https://developer.nytimes.com/get-started>)

2.  Pick an interesting environmental key word(s) and use the jsonlite package to query the API. Pick something high profile enough and over a large enough time frame that your query yields enough articles for an interesting examination.

3.  Recreate the publications per day and word frequency plots using the first paragraph

-   Make some (at least 3) transformations to the corpus (add stopword(s), stem a key term and its variants, remove numbers)

4.  Recreate the publications per day and word frequency plots using the headlines variable (response.docs.headline.main). Compare the distributions of word frequencies between the first paragraph and headlines. Do you see any difference?


#knit to pdf to submit 