---
title: 'topic 3: sentiment analysis i'
author: "Mia Forsline"
date: "4/13/2022"
output: html_document
---
# Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  here,
  LexisNexisTools, #Nexis Uni data wrangling
  lubridate, #working with dates in R 
  pdftools, #read in pdfs 
  sentimentr,
  tidyr, #text analysis in R
  tidytext,
  tidyverse
  )
```

# Introduction 

Sentiment analysis is a tool for assessing the mood of a piece of text. For example, we can use sentiment analysis to understand public perceptions of topics in environmental policy like energy, climate, and conservation. 

This assignment is designed to teach the user more about sentiment analysis using tools such as the [Nexis Uni database through the UCSB Library](https://www.library.ucsb.edu/research/db/211). 

#0. Using the “IPCC” Nexis Uni data set from the class presentation and the pseudo code we discussed, recreate Figure 1A from Froelich et al. (Date x # of 1) positive, 2) negative, 3) neutral headlines):

- read in the IPCC Nexis Uni data set
- extract 100 headlines 
- clean the dataset 

```{r}
my_files <- list.files(pattern = "Nexis_IPCC_Results.docx", 
                       path = here("data"),
                       full.names = TRUE, 
                       recursive = TRUE, 
                       ignore.case = TRUE)

#use lnt_read to read in a LexisNexis file then convert to an object of class 'LNT output' 
dat <- lnt_read(my_files)

meta_df <- dat@meta
articles_df <- dat@articles
paragraphs_df <- dat@paragraphs

dat2 <- data_frame(element_id = seq(1:length(meta_df$Headline)), 
                   Date = meta_df$Date, 
                   Headline = meta_df$Headline) %>% 
        janitor::clean_names()

#remove duplicate headlines, if necessary (in this case, it's not) 
dat2_clean <- dat2[!duplicated(dat2$headline), ]
```

- recreate a similar graph to figure 1A from Froelich et al. 

```{r}
#list of characters
mytext <- get_sentences(dat2$headline)

#sentiments from the text 
sent <- sentiment(mytext)

#create dataframe of sentiments of each sentence 
sentiment_df <- inner_join(dat2, sent, by = "element_id")

sentiment <- sentiment_by(sentiment_df$headline)

#display table of sentences ordered by sentiment 
sentiment_df %>%
  arrange(sentiment)
```
- calculate polarity and assign values of -1, 0, or 1 to indicate negative, neutral, or positive 

```{r}
sentiment_df$polarity <- ifelse(sentiment_df$sentiment <0, -1, ifelse(sentiment_df$sentiment > 0, 1, 0))

#change polarity to an ordered factor 
sentiment_df <- sentiment_df %>% 
  mutate(polarity = factor(polarity, levels = c("1", "0", "-1")))
```

```{r, fig.cap = "Sentiment over time based on headlines with negative (red), neutral (grey), and positive (blue) titles.}
ggplot(data = sentiment_df, aes(x = date, color = polarity)) + 
  stat_count(geom = 'line', aes(y = ..count..)) + 
  labs(y = "Developed Media Sentiment (no. headlines)",
       x = "Date",
       color = "Sentiment",
       title = "IPCC Nexis Uni Data Sentiment from April 4 - 11") + 
  scale_color_manual(values = c("blue", "gray", "red"),
                     labels = c("Positive", "Neutral", "Negative"))
  
```


```{r}
# custom_stop_words <- bind_rows(tibble(word = c("your_word"),  
#                                       lexicon = c("custom")), 
#                                       stop_words)
```


#1. Access the Nexis Uni database through the UCSB library

#2. Choose a key search term or terms to define a set of articles.

I searched for the term "chytrid" from 2020 - 2022 and filtered my results to include only News. 

#3. Use your search term along with appropriate filters to obtain and download a batch of at least 100 full text search results (.docx).

#4. Read your Nexis article document into RStudio.

```{r}
my_files <- list.files(pattern = "chytrid.docx", 
                       path = here("data"),
                       full.names = TRUE, 
                       recursive = TRUE, 
                       ignore.case = TRUE)

#use lnt_read to read in a LexisNexis
file then convert to an object of class 'LNT output' 
dat <- lnt_read(my_files)

meta_df <- dat@meta
articles_df <- dat@articles
paragraphs_df <- dat@paragraphs
```


#5. This time use the full text of the articles for the analysis. First clean any artifacts of the data collection process (hint: this type of thing should be removed: “Apr 04, 2022( Biofuels Digest: http://www.biofuelsdigest.com/ Delivered by Newstex”))

```{r}
#create a dataframe listing all article dates and headlines 
dat2<- data_frame(element_id = seq(1:length(meta_df$Headline)), 
                  Date = meta_df$Date, 
                  Headline = meta_df$Headline) %>% 
  janitor::clean_names()

#remove duplicate headlines (now we have 181 headlines )
dat2 <- dat2[!duplicated(dat2$headline), ]
```


```{r}
#May be of use for assignment: using the full text from the articles
paragraphs_dat <- data_frame(element_id = paragraphs_df$Art_ID, 
                             Text = paragraphs_df$Paragraph)

dat3 <- inner_join(dat2,paragraphs_dat, by = "element_id") %>% 
  janitor::clean_names()

#remove duplicate text rows 
dat3 <- dat3[!duplicated(dat3$text), ]

#remove blank text rows 
dat3 <- subset(dat3, text != " ")

#remove text rows that say "Graphic"
dat3 <- subset(dat3, text != "Graphic")

#remove rows that contain text that indicate an image 
dat3 <- dat3[!grepl(" - ", dat3$text),]

#remove text rows that have tag lines
dat3 <- dat3[!grepl("Apr 0", dat3$text),]

#save text dataframe as CSV 
write_csv(dat3, here::here("data", "chytrid_dat3.csv"))

```

#6. Explore your data a bit and try to replicate some of the analyses above presented in class if you’d like (not necessary).

```{r}

```


#7. Plot the amount of emotion words (the 8 from nrc) as a percentage of all the emotion words used each day (aggregate text from articles published on the same day). How does the distribution of emotion words change over time? Can you think of any reason this would be the case?


We'll start by using the Bing sentiment analysis lexicon.
- Bing lexicon 
```{r get_bing}
bing_sent <- get_sentiments('bing') #grab the bing sentiment lexicon using the tidytext package

#head(bing_sent, n = 20)
```

Here is the starting point for reading in the data as a .csv. We need to unnest the text to the word level so we can label the individual sentiment words. Let's also remove stop words as standard text cleaning procedure. Note: Not every English word is in the lexicons because many English words are pretty neutral.

```{r}
chytrid_text <- read_csv(here::here("data", "chytrid_dat3.csv"))

#unnest to word-level tokens, remove stop words, and join sentiment words
 text_words <- chytrid_text  %>%
  unnest_tokens(output = word, input = text, token = 'words')
 
 sent_words <- text_words%>% #break text into individual words
  anti_join(stop_words, by = 'word') %>% #returns only the rows without stop words
  inner_join(bing_sent, by = 'word') #joins and retains only sentiment words
 
 #head(sent_words) 
```

Create a sentiment score by counting the number of sentiment words occurring per page. We can center the scores around an offset point equal to the average page sentiment score. This lets us measure the sentiment of a given page relative to the overall sentiment of the book.
- in short, is this page more or less pos vs neg ? 

```{r apply_sent_scores}
sent_scores <- sent_words %>%
  count(sentiment, date) %>%
  spread(sentiment, n) %>%
  mutate(raw_score = positive - negative)

ggplot(sent_scores, aes(x = date)) +
  theme_classic() +
  geom_bar(aes(y = raw_score), stat = 'identity', fill = 'slateblue3') +
  #coord_flip() +
  theme(axis.title.y = element_blank()) +
  labs(title = "Sentiment analysis: chytrid",
       y = 'Sentiment score')
         
         
# , #single sentiment score per headline
#   offset = mean(positive - negative), #what is the average sentiment per page? 
#   offset_score = (positive - negative) - offset) %>% #how does this page's sentiment compare to that of the average page?
#   arrange(desc(raw_score))

#sent_scores #sorted by the highest offset_score 


```



##### Origin of the NRC lexicon

"These guys selected about 10,000 words from an existing thesaurus... and then created a set of five questions to ask about each word that would reveal the emotions and polarity associated with it. That's a total of over 50,000 questions.

They then asked these questions to over 2000 people, or Turkers, on Amazon's Mechanical Turk website, paying 4 cents for each set of properly answered questions.

The result is a comprehensive word-emotion lexicon for over 10,000 words."

Let's take a look at the most common sentiment words in the data set

```{r nrc_sentiment}
nrc_sent <- get_sentiments('nrc') #requires downloading a large dataset via prompt
#unique(nrc_sent$sentiment)

#nrc_fear <- get_sentiments("nrc") %>% 
  filter(sentiment == "fear")

#most common words by sentiment
# fear_words <- over_text  %>%
#   unnest_tokens(output = word, input = text, token = 'words') %>%
#   inner_join(nrc_fear) %>%
#   count(word, sort = TRUE)
```

```{r}
nrc_word_counts <- text_words %>%
  inner_join(get_sentiments("nrc")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
```


Let's break it out and plot the contributions by particular words different sentiment categories
```{r sent_counts}
book_sent_counts <- text_words %>%
        group_by(date) %>%
        # mutate(page_num = 1:n(),
        #        index = round(page_num / n(), 2)) %>%
        #unnest_tokens(word, line) %>%
        inner_join(get_sentiments("nrc")) %>%
        group_by(sentiment) %>%
        count(word, sentiment, sort = TRUE) %>%
        ungroup()

book_sent_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

### Introduction to the Nexis Uni data source

```{r nexis_data, message=FALSE}
#setwd("/Users/mateorobbins/Desktop/Git/test/nexis_dat")
#to follow along with this example, download this .docx to your working directory: 
#https://github.com/MaRo406/EDS_231-text-sentiment/blob/main/nexis_dat/Nexis_IPCC_Results.docx

my_files <- list.files(pattern = ".docx", path = getwd(),
                       full.names = TRUE, recursive = TRUE, ignore.case = TRUE)
dat <- lnt_read(my_files) #Object of class 'LNT output'
meta_df <- dat@meta
articles_df <- dat@articles
paragraphs_df <- dat@paragraphs

dat2<- data_frame(element_id = seq(1:length(meta_df$Headline)), Date = meta_df$Date, Headline = meta_df$Headline)

#May be of use for assignment: using the full text from the articles
paragraphs_dat <- data_frame(element_id = paragraphs_df$Art_ID, Text  = paragraphs_df$Paragraph)

dat3 <- inner_join(dat2,paragraphs_dat, by = "element_id")
```

```{r}
#can we create a similar graph to Figure 3A from Froelich et al.? 
mytext <- get_sentences(dat2$Headline)

sent <- sentiment(mytext) #look at the entire headline and tries to correct for negation/context, then gives each headline a rating 

sent_df <- inner_join(dat2, sent, by = "element_id")

sentiment <- sentiment_by(sent_df$Headline)

sent_df %>%
  arrange(sentiment)
```

- calculate mean sentiment per day based on the average sentiment of articles per day 
- plot different lines based on pos/neg 

```{r custom_stop_words}
custom_stop_words <- bind_rows(tibble(word = c("your_word"),  
                                      lexicon = c("custom")), 
                               stop_words)
```



## Assignment (Due 4/19 by 5PM)

1.  Access the Nexis Uni database through the UCSB library: <https://www.library.ucsb.edu/research/db/211>

2.  Choose a key search term or terms to define a set of articles. 

3.  Use your search term along with appropriate filters to obtain and download a batch of at least 100 full text search results (.docx).  

4.  Read your Nexis article document into RStudio. 

5.  This time use the full text of the articles for the analysis.  First clean any artifacts of the data collection process (hint: this type of thing should be removed: "Apr 04, 2022( Biofuels Digest: http://www.biofuelsdigest.com/ Delivered by Newstex"))

- regex 

6.  Explore your data a bit and try to replicate some of the analyses above presented in class if you'd like (not necessary).  

7. Plot the amount of emotion words (the 8 from nrc) as a percentage of all the emotion words used each day (aggregate text from articles published on the same day).  How does the distribution of emotion words change over time?  Can you think of any reason this would be the case?

- line plot 