---
title: 'Topic 7: Word Embeddings'
author: "Mia Forsline"
date: '2022-05-17'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  broom, #creating search_synonym function
  data.table,
  dplyr,
  ggplot2,
  here,
  kableExtra,
  irlba, #singluar value decomposition
  textdata,
  tidytext,
  tidyverse,
  widyr
  )
```

# Assignment: Download a set of pretrained vectors, GloVe, and explore them. 

## Read in the data 
- wrangle to get the dataframe in the proper format to use the synonyms function 
```{r}
glove_data <- fread(here("data", "glove.6B.300d.txt"), header = FALSE) 

glove_df <- glove_data %>%
     remove_rownames() %>%
     column_to_rownames(var = 'V1') #make the first column the index 
```


# 1. Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings. How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

Create the synonyms function 

```{r function}
#take a single word from the word_vectors
#then compare it to the entire matrix
#then output a similarity score

search_synonyms <- function(word_vectors, selected_vector) {
dat <- word_vectors %*% selected_vector
    
similarities <- dat %>%
        tibble(token = rownames(dat), similarity = dat[,1])

similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```

Check similarity scores of words most similar to "fall" and "slip" using the GloVe data 

```{r fall}
glove_matrix <- as.matrix(glove_df)

fall <- search_synonyms(glove_matrix, glove_matrix["fall",])
head(fall, n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "HOLD_position")
```
Compared to the in-class demo, the similarity scores are much higher. Since we are using a different dataset, the word tokens are also no longer specifically climbing related. Instead, the words in the GloVe dataset seem much more general and intuitive. For example, "decline" has the second highest similarity score (after the word "fall" itself) when being compared to the key word of "fall". 

```{r slip}
slip <- search_synonyms(glove_matrix, glove_matrix["slip",])
head(slip, n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "HOLD_position")
```

For "slip," many of the words seem related to the sport of cricket such as "wicket." Highly scored words also include the surnames of famous cricket players such as "dravid." These differences are likely due to us using a completely different set of words compared to the analysis we performed in class. 

```{r plot}
slip %>%
    mutate(selected = "slip") %>%
    bind_rows(fall %>%
                  mutate(selected = "fall")) %>%
    group_by(selected) %>%
    top_n(15, similarity) %>%
    ungroup %>%
    mutate(token = reorder(token, similarity)) %>%
    ggplot(aes(token, similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = "Words",
         y = "Similarity Score",
         title = "What words are most similar to 'slip' or 'fall'?")
         
```
Word math: "snow" and "danger" example 

```{r word math}
snow_danger <- glove_matrix["snow",] + glove_matrix["danger",] 
head(search_synonyms(glove_matrix, snow_danger), n= 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "HOLD_position")
```


```{r}
no_snow_danger <- glove_matrix["danger",] - glove_matrix["snow",] 
head(search_synonyms(glove_matrix, no_snow_danger), n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),                 latex_options = "HOLD_position")
```


# 2. Run the classic word math equation, "king" - "man" = ?

```{r}
king_man <- glove_matrix["king",] - glove_matrix["man",] 
head(search_synonyms(glove_matrix, king_man), n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),                 latex_options = "HOLD_position")
```

# 3. Think of three new word math equations. They can involve any words you'd like, whatever catches you interest.

## a) ball - cricket 
```{r}
no_ball_cricket <- glove_matrix["ball",] - glove_matrix["cricket",] 
head(search_synonyms(glove_matrix, no_ball_cricket), n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),                 latex_options = "HOLD_position")
```

## b) red + apple

```{r}
red_apple <- glove_matrix["red",] + glove_matrix["apple",] 
head(search_synonyms(glove_matrix, red_apple), n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),                 latex_options = "HOLD_position")
```

## c) dog + cat

```{r}
dog_cat <- glove_matrix["dog",] + glove_matrix["cat",] 
head(search_synonyms(glove_matrix, dog_cat), n = 10) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),                 latex_options = "HOLD_position")
```

