---
title: "topic5_hw"
author: "Mia Forsline"
date: '2022-04-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE)

#install packages if necessary, then load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(
  forcats,
  ggraph,
  here,
  igraph,
  kableExtra,
  lubridate,
  pdftools,
  quanteda,
  quanteda.sentiment,
  quanteda.textplots,
  quanteda.textstats,
  readr,
  readtext,
  reshape2,
  sentimentr,
  stringr,
  tidyr,
  tidytext,
  tidyverse,
  widyr,
  wordcloud
  )
```

# Import EPA environmental justice data

```{r}
files <- list.files(path = here::here("data"),
                    pattern = "pdf$", full.names = TRUE)

files <- str_subset(files, pattern="EPA")

ej_reports <- lapply(files, pdf_text)

ej_pdf <- readtext(file = files,
                   docvarsfrom = "filenames",
                   docvarnames = c("type","year"),
                   sep = "_")

#creating an initial corpus containing our data
epa_corp <- corpus(x = ej_pdf, text_field = "text" )
summary(epa_corp)
```

```{r pdf_import}
#I'm adding some additional, context-specific stop words to stop word lexicon
more_stops <-c("2015",
               "2016", 
               "2017", 
               "2018", 
               "2019", 
               "2020", 
               "www.epa.gov", 
               "https")
add_stops <- tibble(word = c(stop_words$word, more_stops)) 
stop_vec <- as_vector(add_stops)
```

Tokenize the data into single words 
```{r}
tokens <- tokens(epa_corp, remove_punct = TRUE)
toks1<- tokens_select(tokens, min_nchar = 3)
toks1 <- tokens_tolower(toks1)
toks1 <- tokens_remove(toks1, pattern = (stop_vec))
dfm <- dfm(toks1)

#first the basic frequency stat
tstat_freq <- textstat_frequency(dfm, n = 5, groups = year)
head(tstat_freq, 10)
```

# 1. What are the most frequent trigrams in the dataset? How does this compare to the most frequent bigrams? Which n-gram seems more informative here, and why?


```{r bigrams}
toks2 <- tokens_ngrams(toks1, n=2)
dfm2 <- dfm(toks2)
dfm2 <- dfm_remove(dfm2, pattern = c(stop_vec))
freq_words2 <- textstat_frequency(dfm2, n=20)
freq_words2$token <- rep("bigram", 20)

tstat_freq2 <- textstat_frequency(dfm2, n = 5, groups = year)
head(tstat_freq2, 10)
```
```{r trigrams}
toks3 <- tokens_ngrams(toks1, n=3)
dfm3 <- dfm(toks3)
dfm3 <- dfm_remove(dfm3, pattern = c(stop_vec))
freq_words3 <- textstat_frequency(dfm3, n=20)
freq_words3$token <- rep("trigram", 20)

tstat_freq3 <- textstat_frequency(dfm3, n = 5, groups = year)
head(tstat_freq3, 10)
```
The third word in the trigrams does not add a lot of value to the n-gram. 

# 2. Choose a new focal term to replace “justice” and recreate the correlation table and network (see corr_paragraphs and corr_network chunks). Explore some of the plotting parameters in the cor_network chunk to see if you can improve the clarity or amount of information your plot conveys. Make sure to use a different color for the ties!

Now we'll create some different data objects that will set us up for the subsequent analyses

```{r tidy}
#convert to tidy format and apply my stop words
raw_text <- tidy(epa_corp)

#Distribution of most frequent words across documents
raw_words <- raw_text %>%
  mutate(year = as.factor(year)) %>%
  unnest_tokens(word, text) %>%
  anti_join(add_stops, by = 'word') %>%
  count(year, word, sort = TRUE)

#number of total words by document  
total_words <- raw_words %>% 
  group_by(year) %>% 
  summarize(total = sum(n))

report_words <- left_join(raw_words, total_words)
 
#paragraph tokens 
par_tokens <- unnest_tokens(raw_text, output = paragraphs, input = text, token = "paragraphs")

#give each paragraph an id number 
par_tokens <- par_tokens %>%
 mutate(par_id = 1:n())

#individual words 
par_words <- unnest_tokens(par_tokens, output = word, input = paragraphs, token = "words")

```

Let's see which words tend to occur close together in the text. This is a way to leverage word relationships (in this case, co-occurence in a single paragraph) to give us some understanding of the things discussed in the documents.

```{r co-occur_paragraphs}
word_pairs <- par_words %>% 
  pairwise_count(word, par_id, sort = TRUE, upper = FALSE) %>%
  anti_join(add_stops, by = c("item1" = "word")) %>%
  anti_join(add_stops, by = c("item2" = "word"))

word_pairs %>%
  filter(n >= 70) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "dodgerblue") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()
```

```{r corr_paragraphs}
#calculate correlation coefficients between words 
word_cors <- par_words %>% 
  add_count(par_id) %>% 
  filter(n >= 50) %>% 
  select(-n) %>%
  pairwise_cor(word, par_id, sort = TRUE)

#search for words correlated with "justice" 
urban_cors <- word_cors %>% 
  filter(item1 == "urban")

#graph correlation coefficients for words correlated with 4 key terms 
  word_cors %>%
  filter(item1 %in% c("waters", 
                      "flooding", 
                      "housing", 
                      "urban"))%>%
  group_by(item1) %>%
  top_n(6) %>%
  ungroup() %>%
  mutate(item1 = as.factor(item1),
  name = reorder_within(item2, correlation, item1)) %>%
  ggplot(aes(y = name, x = correlation, fill = item1)) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~item1, ncol = 2, scales = "free")+
  scale_y_reordered() +
  labs(y = NULL,
         x = NULL,
         title = "Correlations with key words",
         subtitle = "EPA EJ Reports")
  
#let's zoom in on just one of our key terms
urban_cors <- word_cors %>% 
  filter(item1 == "urban") %>%
  mutate(n = 1:n())
```

```{r corr_network}
urban_cors  %>%
  filter(n <= 50) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation, edge_width = correlation), edge_colour = "cyan4") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()
```

# 3. Write a function that allows you to conduct a keyness analysis to compare two individual EPA reports (hint: that means target and reference need to both be individual reports). Run the function on 3 pairs of reports, generating 3 keyness plots.

```{r}
keyness <- textstat_keyness(dfm, target = 1) #target = 1 --> first document in the list, so in this case 2015
#target = 2 would use 2016 as the target document
textplot_keyness(keyness)

#chi-square values for each word 
#in 2015, climate has the highest chi-square value, so climate occurred the most "more than expected" given the baseline frequency 
#words in blue = occured more than expected
#words in grey = occurred less than expected 
#reference = includes all reports except 2015 
```

```{r}


keyness_function <- function(report_number) {
  
  for (i in (1:(length(epa_corp) -1))) {
    report_test <- epa_corp[i: (i+1)]
    print(report_test)
    tokens_test <- tokens(report_test, remove_punct = TRUE) #list split word by word
    tokens_test <- tokens_select(tokens_test, min_nchar = 3)
    tokens_test <- tokens_tolower(tokens_test)
    tokens_test <- tokens_remove(tokens_test, pattern = (stop_vec))
    dfm_test <- dfm(tokens_test) #document feature matrix

    keyness <- textstat_keyness(dfm_test, target = 1)
    print(textplot_keyness(keyness))
    
  }
  
}

keyness_function(report_number = 1)



```



# 4. Select a word or multi-word term of interest and identify words related to it using windowing and keyness comparison. To do this you will create to objects: one containing all words occurring within a 10-word window of your term of interest, and the second object containing all other words. Then run a keyness comparison on these objects. Which one is the target, and which the reference? Hint

